Prediction of 2024 US election ...

### 2024-11-01 paper/paper.qmd
update paper

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
library(arrow)
library(rstanarm)
library(brms)
```

@@

@@

```{r}
#| label: fig-clean-data-histogram-Harris-Trump
#| fig-cap: the average PCT vs State for Harris and Trump
#| echo: false
#| warning: false
#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
#| layout-nrow: 2
#| fig-width: 40
#| fig-height: 20
Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
Harris_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = avg_pct)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
Trump_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = avg_pct)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

@@

@@

```{r}

# Transform Biden and Trump’s vote shares to fit a beta distribution
# Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
total_votes <- 100
biden_shape1 <- 51.3 / total_votes * 10
biden_shape2 <- (1 - 51.3 / total_votes) * 10
trump_shape1 <- 46.8 / total_votes * 10
trump_shape2 <- (1 - 46.8 / total_votes) * 10

# Set up priors using beta distributions based on 2020 vote shares
# prior <- c(
#   set_prior(paste0("beta(", biden_shape1, ", ", biden_shape2, ")"), class = "b", coef = "biden"),
#   set_prior(paste0("beta(", trump_shape1, ", ", trump_shape2, ")"), class = "b", coef = "trump")
# )

prior <- c(
  set_prior(beta(biden_shape1, biden_shape2), class = "b", coef = "biden"),
  set_prior(beta(trump_shape1, trump_shape2), class = "b", coef = "trump"))

# Define the model formula

formula1 <- pct ~ pollscore + days_taken_from_election + methodology

# Run the Bayesian model
model <- brm(
  formula = formula1,
  data = baye_model_data,
  family = categorical(),
  prior = prior,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  cores =4
)

# Summarize the model
summary(model)

>>>>

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
library(arrow)
library(rstanarm)
library(brms)
library(dplyr)
library(lubridate)
```

@@

The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here] and lubridate [@lubridate].

@@

```{r}
#| label: fig-clean-data-histogram-Harris
#| fig-cap: the average PCT vs State for Harris
#| echo: false
#| warning: false
#| height: 40

Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
Harris_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = avg_pct, y = state)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

@@

```{r}
#| label: fig-clean-data-histogram-trump
#| fig-cap: the average PCT vs State for Trump
#| echo: false
#| warning: false
#| height: 40

Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
Trump_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = avg_pct, y = state)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

@@
